\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{courier}
\usepackage{fancyhdr}
\usepackage{color}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{graphicx}


\definecolor{light-gray}{gray}{0.95}

\lstset{
 showspaces=false,
 showtabs=false,
 frame=none,
 tabsize=2,
 breaklines=true,
 numbers=none,
 showstringspaces=false,
 breakatwhitespace=true,
 escapeinside={(*@}{@*)},
 keywordstyle=\bfseries,
 basicstyle=\footnotesize\ttfamily,
}

\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}

\setlength\parindent{0pt} % Indentazione paragrafi
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex} % Spaziatura paragrafi

\title{SDN controller clustering \\ \large Computer Networks module - SDN assignment}
\author{Michele Zanotti}
\date{Spring term 2018}

\pagestyle{fancy}
\fancyhf{}
\lhead{Computer Networks}
\rhead{SDN assignment}
\rfoot{\thepage}

\begin{document}

\maketitle
\newpage

\section*{Overview}
In this paper are proposed four different lab activities meant to teach how to implement
a cluster of controllers inside Mininet using some of the most common available methods.
Each activity focuses on a single method and use it to implement a certain topology.

In particular, in each activity the following methods will be explained:
\begin{itemize}
  \item \textbf{Activity 1}: implement a network with multiple local controllers using a python
  script and the middle-level Mininet API.
  \item \textbf{Activity 2}: implement a network with multiple local controllers using a python
  script, a custom switch class and the topology classes provided by the Mininet high-level API.
  \item \textbf{Activity 3}: implement a network with multiple controllers using
  \emph{miniedit}, a tool provided by Mininet.
\end{itemize}

The paper also includes a fourth lab activity (\textbf{Activity 4}), which is a challange meant
to let the reader test the knowledge acquired with the execution of the previous
four activities.

It's important to point out that all the the proposed activities are meant to be
only an introduction to SDN controller clustering, therefore this paper won't cover
some of the aspects that has to be considered when implementing a cluster of
controllers in a real-world context, such as load balancing and communication
between controllers. More information about these topics can be found in
\cite{ref-1}, \cite{ref-2} and [].

The communication between different controllers can however be automatically managed
by Mininet, the network emulator used in this paper activities. When in fact a
network with multiple local controllers is implemented inside the emulator, all
the local controllers are executed inside the kernel space on a single machine
and therefore they can communicate with each other without an external
communication mechanism being necessary. [ REFERENCE ???]

\include{activity-1}
\include{activity-2}


\begin{thebibliography}{99}
  \footnotesize

  \bibitem{ref-1}
  N. T. Hai and D. S. Kim,
  ``Efficient load balancing for multi-controller in SDN-based mission-critical networks'',
  \textit{2016 IEEE 14th International Conference on Industrial Informatics (INDIN)},
  Poitiers,
  2016,
  pp. 420-425.

  \bibitem{ref-2}
  K. Phemius, M. Bouet and J. Leguay,
  ``DISCO: Distributed SDN controllers in a multi-domain environment'',
  \textit{2014 IEEE Network Operations and Management Symposium (NOMS)},
  Krakow,
  2014,
  pp. 1-2.

  \bibitem{ref-3}
  GitHub. (2016).
  \textit{mininet}. [online].
  Available at: \url{https://github.com/mininet/mininet/blob/master/examples/controllers2.py}
  [Accessed 28 Mar. 2018].
\end{thebibliography}

\end{document}
