\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{courier}
\usepackage{fancyhdr}
\usepackage{color}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{url}
\usepackage{textcomp}
\usepackage{pdfpages}
\usepackage[colorlinks]{hyperref}
\hypersetup{urlcolor=blue}




\setlength\parindent{0pt} % Indentazione paragrafi
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex} % Spaziatura paragrafi

\begin{document}







\clearpage
\section*{List of Abbreviations}
\begin{tabular}{ p{4cm} l }
 3GPP & Third Generation Partnership Project \\
 CN & Core network \\
 CSG & Closed subscriber group (of a home eNodeB) \\
 E-UTRAN & Evolved UMTS terrestrial radio access network \\
 EPS & Evolved packet system \\
 GGSN & Gateway GPRS Support Node \\
 GTP & GPRS tunnelling protocol \\
 GTP-C & GPRS tunnelling protocol control part \\
 GTP-U & GPRS tunnelling protocol user part \\
 IMEI & International mobile equipment identity \\
 LTE & Long term evolution \\
 MME & Mobility management entity \\
 NSAPI & Network Service Access Point Identifier \\
 PDN & Packet data network \\
 PDU & Protocol data unit \\
 P-GW & Packet data network gateway \\
 RAB & Radio access bearer \\
 RNC & Radio network controller \\
 SDU & Service data unit \\
 SGSN & Serving GPRS Support Node \\
 S-GW & Serving gateway \\
 S-RNTI & Serving RNC Radio Network Temporary Identifier \\
 TEID & Tunnel endpoint identifier \\
 UE & User equipment \\
 UMTS & Universal Mobile Telecommunications System \\
 UTRAN & UMTS terrestrial radio access network \\


\end{tabular}

\clearpage
\section{Introduction}
The intra-RAT handover from E-UTRAN to UTRAN is composed of two phases, each of
which involves (as well as the UE) either nodes from the LTE network and nodes
from the UMTS netowork: the \emph{preparation phase} and the \emph{execution phase}.
In particular, the LTE elements involved in the procedure are:
\begin{itemize}
  \item eNodeB
  \item E-UTRAN network
  \item MME
  \item S-GW
  \item P-GW
\end{itemize}
while the UMTS elements are:
\begin{itemize}
  \item RNC
  \item SGSN
  \item GGSN
\end{itemize}

It is assumed that both the LTE and the UMTS networks belong to the same
operator. This means that the P-GW of the LTE network and the GGSN of the UMTS
network are the same physical device, therefore, since in this paper the focus
is on the LTE network, only the P-GW will be showed. Moreover, since the goal
of this paper is to show only the intra-RAT handover procedure, some non-central
aspects such as the relocation of the S-GW will not be considered and the related
messages and parameters will be left out.

It is important to point out that before, during and after the handover the UE
is in the state ``LTE\_ACTIVE'' and user data is uploaded/downloaded (otherwise
if the UE was in idle state it wouldn't be a handover but a simple cell reselection).
In particular, as shown in figure \ref{fig:pre-handover-links}, before the handover procedure both uplink
and downlink user data is transmitted through:
\begin{itemize}
	\item bearer(s) between UE and source eNodeB
	\item GTP\footnote{GPRS Tunneling protocol (GTP) is a IP/UDP based protocol used to
	encapsulate user data when passing through core network (GTP-U) or to carry bearer
	specific signalling traffic between various core network nodes (GTP-C).}
  tunnel(s) between	source eNodeB, S-GW and P-GW.
\end{itemize}


The figure \ref{fig:architecture-model} shows the elements involved in this
handover procedure, how they are connected to each other and through which
interfaces, figure \ref{fig:pre-handover-links} shows how data is transmitted
before the start of the procedure.



\include{preparation}
\include{execution}
\include{reject}


\begin{thebibliography}{99}
\footnotesize

  \bibitem{direct-tunnelling}
  Cisco. (2017).
  \textit{P-GW Administration Guide, StarOS Release 20 - Direct Tunnel for 4G (LTE) Networks [Cisco ASR 5000 Series]}. [online]
  Available at: \url{https://www.cisco.com/c/en/us/td/docs/wireless/asr_5000/20/P-GW/b_20_PGW_Admin/b_20_PGW_Admin_chapter_011111.html}
  [Accessed 5 May 2018].

  \bibitem{routing-area-update}
  3GPP TS 23.401 (September 2011)
  \textit{General Packet Radio Service (GPRS) Enhancements for Evolved Universal
  Terrestrial Radio Access Network (E-UTRAN) Access},
  Release 10,
  sections 5.3.3.3, 5.3.3.6.

  \bibitem{channels}
  Poole, I. (n.d.).
  \textit{LTE Physical, Logical and Transport Channels :: Radio-Electronics.Com}. [online]
  Radio-electronics.com.
  Available at: \url{http://www.radio-electronics.com/info/cellulartelecomms/lte-long-term-evolution/physical-logical-transport-channels.php}
  [Accessed 16 May 2018].


\end{thebibliography}


\end{document}
